---
title: "hamadryas_beh"
output: html_document
date: '2022-07-28'
---

# DATA LOGGED SINCE 24 AUG
```{r}
data <- read.csv("zm_final.csv")

library(vegan)
library(ggplot2)
library(lme4)
library(tidyverse)
library(mvabund)
library(dplyr)

# CLEANING UP OF DATASET

data2 <- data[ , c(1, 9, 21, 23, 25, 28, 30, 32:36, 38:39, 41:44)]

library(memisc)

data2 <- memisc::rename(data2, ## rename() exists in multiple packages. need to specify.
                         All.Occurrence.Value = "AllOcc.Beh",
                         All.Occurrence.Behavior.Social.Modifier = "AllOcc.Recipient",
                         Interval.Channel.1.Value = "Int.Beh",
                         Interval.Channel.1.Behavior.Modifier.1 = "Role",
                         Interval.Channel.1.Social.Modifier = "Int.Recipient",
                         Space.Use.Grid.X = "Grid.X",
                         Space.Use.Grid.Y = "Grid.Y",
                         Space.Use.Coordinate.X = "Coord.X",
                         Space.Use.Coordinate.Y = "Coord.Y",
                         Question..No..of.adult.males.present..total.1_69 = "M.Present",
                         Question..No..of.adult.males.sterilised_71 = "M.Sterile",
                         Question..No..of.adult.females.present..total_70 = "F.Present",
                         Question..No..of.adult.females.sterilised_72 = "F.Sterile",
                         warn = FALSE
                         )

library(formattable)

data2$PercentSM <- data2$M.Sterile / data2$M.Present
data2$PercentSM <- formattable::percent(data2$PercentSM)

data2$PercentSF <- data2$F.Sterile / data2$F.Present
data2$PercentSF <- formattable::percent(data2$PercentSF)
```


# ACTIVITY BUDGET PLOTS
```{r}
# overall (all males)

data.all <- as.data.frame(table(data2$Int.Beh))

data.all <- filter(data.all, Var1 != '') #remove empty rows where no behaviour recorded (forgot to record?)

beh.vec <- c("Affiliative", "Aggressive", "Allogroom", "Dominance", "Submissive", "Vigilance", "Locomotion", "Forage", "Self-directed", "Rest", "Sleep", "Other", "Behaviour unclear")

data.all <- data.all %>% mutate(Var1 = factor(Var1, levels = beh.vec)) %>%
  arrange(Var1) # rearranging Var1

#create column of short forms, then rearrange similarly. using mutate again will mess up dataset, so just use factor.

beh.vec.short <- c("Aff", "Agg", "Allo", "Dom", "Sub", "Vigil", "Loc", "Forage", "Self", "Rest", "Sleep", "Other", "Unclear")

data.all$shortForm <- beh.vec.short

data.all$shortForm <- factor(data.all$shortForm, levels = beh.vec.short)
```


```{r}
ggplot(data.all, aes(shortForm, Freq, fill = Var1)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  theme_classic() +
    labs(x = "Behaviour", y = "Counts", fill = "Behaviour")
```

```{r}
#individual males

data.indiv <- data2 %>% dplyr::count(Focal.Name, Int.Beh)

data.indiv <- filter(data.indiv, Int.Beh != '')
```

```{r}
ggplot(data.indiv, aes(Int.Beh,n, fill = Int.Beh)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  theme_classic() +
    labs(x = "Behaviour", y = "Counts", fill = "Behaviour") +
      facet_wrap(.~Focal.Name)
```



# CREATING EDGE LISTS - SNA
```{r}
library(igraph)
library(igraphdata)
library(qgraph)
library(kableExtra)

aff.vec <- c("Gential touch", "Notifying present", "Lipsmack")
agg.vec <- c("Threat", "Lunge", "Chase", "Contact aggression", "Neck bite", "Interposition")
males.vec <- c("Bear", "Blondie", "Boss", "Helmet", "Jeremy", "King", "Mascot", "Poodle", "Rambutan", "Romeo", "Scruffy", "Totem", "Uncle", "Zzu", "Ralph")

full.data <- data2 %>% dplyr::select(Focal.Name, AllOcc.Beh, AllOcc.Recipient) %>%
  memisc::rename(Focal.Name = source, AllOcc.Beh = behaviour, AllOcc.Recipient = target) %>%
  filter(source %in% males.vec & target %in% males.vec)

aff.data <- full.data %>% filter(behaviour %in% aff.vec)
agg.data <- full.data %>% filter(behaviour %in% agg.vec)

# CREATING EDGE DF

aff.edges <- aff.data %>% group_by(source) %>%
  count(target) %>%
  memisc::rename(n = weight)

agg.edges <- agg.data %>% group_by(source) %>%
  count(target) %>%
  memisc::rename(n = weight)

# CREATING NODE DF

#aff.nodes <- aff.data %>% dplyr::select(source) %>%
  #group_by(source) %>%
 # summarise() %>%
 # rename(source = name)

#nodes <- as.data.frame(males.vec)
```



# SNA NETWORKS, +ve KPA
```{r}
aff.qgraph = qgraph(aff.edges,
                    edge.labels = TRUE, # FALSE to remove numbers
                    edge.color = "blue",
                    title = "Affiliative relationships")
```

```{r}
# AFF NETWORK +ve KPA

aff.igraph <- graph_from_data_frame(aff.edges, directed = TRUE, vertices = NULL)
#create an igraph graph from aff.edges (cannot use the one from qgraph cause diff package)

library(influenceR)

set.seed(1) # set a random seed generator to reproduce your analysis
keyplayer(aff.igraph, k = 1) # k = number of key players

```


```{r}
agg.qgraph = qgraph(agg.edges,
                    edge.labels = TRUE,
                    edge.color = "red",
                    title = "Aggressive relationships")
```

affiliative: more reciprocal relationships
- most of the agg interactions are "chase", which is one-way. threats could be exchanged before chase commences but missed out while sampling
 
ralph only receives aggression (does not return)

```{r}
# AGG NETWORK +ve KPA

agg.igraph <- graph_from_data_frame(agg.edges, directed = TRUE, vertices = NULL)
#create an igraph graph from aff.edges (cannot use the one from qgraph cause diff package)
#agg.igraph <- as.igraph.qgraph(agg.graph, attributes = TRUE)

library(influenceR)

set.seed(1)
keyplayer(agg.igraph, k = 1)
```


# converting graphs to adjacency matrices
```{r}
#library(keyplayer)
#library(sna)

aff.adj <- get.adjacency(aff.igraph,
              names = TRUE,                   #vertice names
              attr = 'weight',                #for weighted adjmat
              sparse = FALSE)
#convert graph to adjacency matrix

agg.adj <- get.adjacency(agg.igraph,
              names = TRUE,                   #vertice names
              attr = 'weight',                #for weighted adjmat
              sparse = FALSE)

#write.csv(aff.adj, "aff_adj.csv", row.names = FALSE)
#write.csv(agg.adj, "agg_adj.csv", row.names = FALSE)
```


# individual degree centrality
```{r}
central.aff <- centralityTable(aff.qgraph)

central.agg <- centralityTable(agg.qgraph)

# outdegree: outgoing connections
# indegree: incoming connections
# closeness: a node’s capability of quickly reaching other nodes
# betweenness: a node’s brokerage power in a network
```

```{r}
#transforming weird table to useful df

central.aff.df <- data.frame()

central.aff.df <- central.aff[1:15,]
central.aff.df <- select(central.aff.df, node)

list.col <- list()

for (i in unique(central.aff$measure)) {
  c <- central.aff %>%
    filter(measure == i) %>%
    dplyr::select(value)
  list.col[[i]] <- c
}

for (i in list.col) {
  central.aff.df <- append(central.aff.df, i)
}

names(central.aff.df) <- c("node", "Betweenness", "Closenness", "InStrength", "OutStrength", "OutExpectedInfluence", "InExpectedInfluence")

```


# detection of communities (based off affiliative SNA)
```{r}
aff.igraph.undir <- as.undirected(aff.igraph, mode = 'collapse')
# makes graph undirected for comm det code to work

aff.louvain <- cluster_louvain(aff.igraph.undir)

aff.louvain.member <- data.frame(node = 1:gorder(aff.igraph.undir), community = aff.louvain$membership)

table(aff.louvain$membership) # to see the number of communities and the number of nodes in each community

modularity(aff.louvain) # computes modularity
```
```{r}
par(mar=c(0,0,0,0)+.1)
set.seed(1)

V(aff.igraph.undir)$community <- aff.louvain$membership

plot(aff.igraph.undir,
     vertex.color = V(aff.igraph.undir)$community,
     vertex.label = V(aff.igraph.undir)$community,
     label.color = 'black',
     edge.color = 'darkgrey',
     vertex.size = 10,
     vertex.frame.color = 'white')
```



# DISTANCE
```{r}
data.dist <- filter(data2, Channel.Type == 'Interval')
data.dist$newSessID <- paste(data.dist$SessionID, data.dist$Frame.Number, sep = "_")

list.1 <- list() ## create new list

library(matrixcalc)

for (i in unique(data.dist$newSessID)) { ## iterate through SessionID

  ## create df of names and coordinates for session 
  df_coord <- data.dist %>% filter(newSessID == i) %>%
    dplyr::select(Focal.Name, Coord.X, Coord.Y, newSessID)
  
  rownames(df_coord) <- df_coord$Focal.Name ## to ensure matrix rows and columns are labelled with names
  
  mat_dist <- as.matrix(dist(df_coord[2:3], method = "euclidean")) ## dist() produces a class, hence as.matrix() is required
  ## create matrix of distances, using only coordinate columns in df (hence [2:3])
  
  mat_dist <- lower.triangle(mat_dist) ## replaces values in upper triangle of matrix with 0

  df_dist <- as.data.frame(mat_dist, rownames = TRUE) ## transform matrix into df
  
  df_dist[df_dist == 0] <- NA ## change 0 to NA
  
  df_dist$SessionID <- df_coord$newSessID ## add SessionID column
  
  list.1[[i]] <- df_dist ## add df to list
}
```


# merging all df in list
```{r}
library(reshape)

merged.lists <- list.1 %>% reduce(full_join)

# creating column for names since rownames() does not allow duplicate row names. (everything is still in order!
# rows in merged are currently numbered.
merged.lists$FocalMale <- data.dist$Focal.Name

# moving SessionID and FocalMale columns to left of df
merged.lists <- merged.lists %>% relocate(SessionID, .before = Bear) %>%
  relocate(FocalMale, .after = SessionID)

```


# finding mean and sd of pairwise distances
```{r}
library(tidyselect)

pos_mean <- as.data.frame(   ## because this produces a tibble, which rownames() does not work on
  merged.lists %>% group_by(FocalMale) %>%
  summarise(across(Bear:Uncle, ~mean(.x, na.rm = TRUE), '{.col}')))

rownames(pos_mean) <- pos_mean$FocalMale

pos_mean <- pos_mean %>% .[-1] %>%    ## remove FocalMale col
  dplyr::select(sort(peek_vars()))       ## sort columns by alphabetical order

pos_sd <- as.data.frame(
  merged.lists %>% group_by(FocalMale) %>%
  summarise(across(Bear:Uncle, ~sd(.x, na.rm = TRUE), '{.col}')))

rownames(pos_sd) <- pos_sd$FocalMale

pos_sd <- pos_sd %>% .[-1] %>%
  dplyr::select(sort(peek_vars()))

# across works within dplyr verbs like summarise. '{col}' at the end retains names of original columns. https://dplyr.tidyverse.org/reference/across.html
```


# rescaling
```{r}
library(scales)

rescaled_dist <- merged.lists %>% group_by(FocalMale) %>%
  summarise(
      suppressWarnings(  ## must be placed directly in front of the fn that is producing warnings
        across(Bear:Uncle,
               ~rescale(.x,
                        to = c(0, 100),
                        from = range(.x, na.rm = TRUE, finite = TRUE)))))

rescaled_dist <- rescaled_dist[rowSums(is.na(rescaled_dist)) != ncol(rescaled_dist[2:15]),]
#remove all rows with only NaN values
```
 

# sna graph for distance
```{r}
# convert df of mean dist -> sna edge table of source, target, weight
pos_mean <- as.matrix(pos_mean)
dist.edges <- reshape::melt(pos_mean)

dist.edges <- na.omit(dist.edges) # removing rows with NaN values

# taking top 25% (22 of 91) values
library(data.table)
dist.edges <- data.table(dist.edges, key = "value") #sorts "value" col in ascending order
top25 <- dist.edges[, head(.SD, 22)]

#bottom 25%
bot25 <- dist.edges[, tail(.SD, 22)]

qgraph(top25,
       directed = FALSE,
       edge.labels = FALSE,
       edge.color = "green",
       title = "Shortest 25% of pairwise distances")
```
```{r}
qgraph(bot25,
       directed = FALSE,
       edge.labels = FALSE,
       edge.color = "green",
       title = "Longest 25% of pairwise distances")
```


# UTILISATION DISTRIBUTION
```{r}
library(amt)

```

# graphs of individual space use
```{r}
data2$newdate = as.POSIXct(data2$DateTime) #create new column for datetime in correct format

data2$day.no = lubridate::yday(data2$newdate) #yday returns the day of the year as a number

# every graph is an individual -- data points are entries across all sessions
ggplot(data2, aes(x = Coord.X, y = Coord.Y, colour = day.no)) +
  geom_point(size = 1.5) +
  facet_wrap(.~Focal.Name)
```
